---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

# Course project 1
===========================
Load libraries
```{r message=FALSE}
library(tidyverse)
library(ggplot2)
```
1. Code for reading in the dataset and/or processing the data
```{r,echo =TRUE, message=FALSE}
activity<-read.table("activity.csv", header = TRUE, sep=",")
head(activity)
```
2. Histogram of the total number of steps taken each day
Total steps for each day
```{r}
activity %>% group_by(date) %>% summarise(steps_day = sum(steps, na.rm = TRUE)) %>%  ggplot(aes(x = steps_day)) + geom_histogram(bins = 20) + ggtitle("Histogram of the total number of steps taken each day") + xlab("Steps Per Day") + ylab("Frequency")
```

3. Mean and median number of steps taken each day
Mean number of steps taken each day
```{r}
mean_day<-activity %>% group_by(date) %>% summarise(mean_day = mean(steps, na.rm = TRUE))
```
Mean of steps each day
```{r}
mean_day %>% summarise(total_mean = mean(mean_day, na.rm=TRUE))
```

Meadian number of steps taken each day
```{r}
median_day<-activity %>% group_by(date) %>% summarise(median_day = median(steps, na.rm = TRUE))
```

Median of steps each day
```{r}
median_day %>% summarise(total_mean = mean(median_day, na.rm=TRUE))
```

4. Time series plot of the average number of steps taken

Change class date 
```{r}
z<-mean_day$date <- as.Date(mean_day$date)
mean_day<-mutate(mean_day, date=z)
```


```{r}
mean_day %>%  ggplot(aes(x = date, y = mean_day)) + geom_point() + geom_line(color="maroon") + ggtitle("Mean of steps taken each day") 
```

5.The 5-minute interval that, on average, contains the maximum number of steps
Group by interval, calculate mean of steps in each interval and filter by the max value of steps in the interval
```{r}
a<- activity %>% group_by(interval) %>% summarise(step_interval = mean(steps, na.rm = TRUE)) %>% filter(step_interval==max(step_interval))
```
```{r}
print(a)
```

```{r}
paste("The 5-minute interval that, on average, contains the maximum number of steps is", a$interval)
```

6. Code to describe and show a strategy for imputing missing data
How many NA values are in the data
```{r}
sum(is.na(activity))
```

Import data eliminating missing values
```{r}
activity_no_NA <- activity[complete.cases(activity),]
```

Corroborate no NA values in activity_no_NA
```{r}
sum(is.na(activity_no_NA))
```
7.Histogram of the total number of steps taken each day after missing values are imputed
```{r}
activity_no_NA %>% group_by(date) %>% summarise(steps_day = sum(steps, na.rm = TRUE)) %>%  ggplot(aes(x = steps_day)) + geom_histogram(bins = 20) + ggtitle("Histogram of the total number of steps taken each day") + xlab("Steps Per Day") + ylab("Frequency")
```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

Define day of the week
```{r}
y<-activity_no_NA%>%mutate(week = weekdays(as.Date(date)))
```

Define weekdays and weekends
```{r}
z<-mutate(y, weekday_weekend = ifelse(week=="Saturday"| week == "Sunday", "weekend", "weekday"))
```

```{r}
mean_interval<-z%>%group_by(weekday_weekend, interval)%>%summarise(mean_interval=mean(steps))
ggplot(mean_interval, aes(x=interval, y=mean_interval))+geom_line()+facet_wrap(~weekday_weekend)+ylab("Mean steps")+xlab("Interval")+ggtitle("Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends")
```

